---
title: "Game 4: Shapiro vs Leopold Le Ruyet"
author: "Jay Shapiro"
date: "4/xx/2020"
output: 
  html_document:
    theme: cerulean
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>

You know the drill by now.

<br>

1 **e4 e5**<br>
2 **Nf3 Nc6**<br>
3 **Bb5 a6**<br>
4 **Ba4 Nf6**<br>
5 **0-0 Be7**<br>
6 **Re1 b5**<br>
7 **Bb3 0-0**<br>
8 **c3 d5**<br>

```{r echo = FALSE, warning = FALSE, message = FALSE, include = FALSE}
library(rchess)
library(tidyverse)
library(ggplot2)
library(plotly)

game <- Chess$new()

fen <- "r1bq1rk1/2p1bppp/p1n2n2/1p1pp3/4P3/1BP2N2/PP1P1PPP/RNBQR1K1 w - - 0 9"

game$load(fen)
```
```{r echo = FALSE, warning = FALSE, message = FALSE}

plot(game)

```

<br>

The move 8...d5 signals the Marshall Attack, an enterprising gambit in the Ruy Lopez where Black forfeits a central pawn for a promising kingside attack. The mainline theory runs very deep in the Marshall, with the lichess.org database having a whopping 934 games played from the position after White's sixteenth move!

<br>

9 **exd5 Nxd5**<br>
10 **Nxe5 Nxe5**<br>
11 **Rxe5 c6**<br>
12 **d4 Bd6**<br>
13 **Re1 Qh4**<br>
14 **g3 Qh3**<br>
15 **Be3 Bg4**<br>
16 **Qd3 f5**<br>

```{r echo = FALSE, warning = FALSE, message = FALSE, include = FALSE}
fen <- "r4rk1/6pp/p1pb4/1p1n1p2/3P2b1/1BPQB1Pq/PP3P1P/RN2R1K1 w - - 0 17"

game$load(fen)

```

```{r echo = FALSE, warning = FALSE, message = FALSE}

plot(game)

```

<br>

Finally deviating from the main line, which is 16...Rae8, and choosing a move which is far less popular but still theory

<br>

17 **f4 Rae8**<br>
18 **Nd2 Kh8**<b>
19 **Bxd5 cxd5**<br>
20 **a4 bxa4**<br>
21 **Qxa6 Rxe3!**<br>

This move surprised me; I was expecting the more natural Re6, but both moves are playable and have been played before.

<br>

22 **Rxe3 Bxf4**<br>
23 **Qd3?! ...**<br>

Finally deviating from previously played games (23 Nf1) and striking out into uncharted territory. Although Qd3 is objectively worse than Nf1, I had overweening ambitions about my prospects this game, and I was willing to accept a worse position that allowed me to still push for a win.

<br>

23 **... Bxe3+**<br>
24 **Qxe3 f4**<br>
25 **Qe7 Kg8**<br>
26 **Rf1 f3**<br>
27 **Rf2 h5?**<br>

Black makes the mistake of boxing his own queen into a claustrophobia-inducing prison on h3. Black's best option would have been to extricate his piece with Qh5 and give up on the pipe dream of mating on g2.

<br>

28 **Qe5 h4**<br>
29 **gxh4? ...**<br>

The simple 29 Qxd5+ Kh8 30 Qe5 was superior for White.

<br>

29 **... Rf6?**<br>

Again, the simpler move 29...Qxh4 would have been stronger. Black makes matters needlessly complicated.

<br>

30 **Qg3 ...**<br>

Still ignoring the pawn on d5 in favor of the king safety, and also trying to bait Black into playing that move that he does. I thought it was very likely he would automatically play it and overlook the refutation.

<br>

30 **... Rg6??**<br>
31 **Nxf3 Bxf3??**<br>

I was expecting him to play 31...Bf5 and pin my queen to my king, which I had planned to meet with 32 Ng5, to break the pin. I wasn't expecting him to just give up the rook on g6 and allow me to simplify into an easily winning endgame.

<br>

32 **Qxg6 Be4**<br>
33 **Qe8+ Kh7**<br>
34 **Qh5+ Kg8**<br>
35 **Qe8+ Kh7**<br>
36 **Qxa4 ...**<br>

Covering the d1 square to avoid a perpetual check from Black (Qg4-Qd1-Qg4) and picking up a significant material advantage. Black will win the Exchange back over the next few moves, but White will still be ahead by a couple of pawns.

<br>

36 **... Qg4+**<br>
37 **Kf1 Bd3+**<br>
38 **Ke1 Qg1+**<br>
39 **Kd2 Qxf2+**<br>
40 **Kxd3 Qxb2**<br>

Winning from this point is just simple technique. White needs to march his king up the board while fending off checks from the desperate Black queen.

<br>

41 **Qc2 Qa1**<br>
42 **c4 Qf1+**<br>
43 **Kc3+ Kh8**<br>

Picking up another free pawn, White will soon be an extra queen to the good.

<br>


44 **cxd5 Qa1+**<br>
45 **Kd3 Qa6**<br>
46 **Qc4 Qg6+**<br>
47 **Kc3 Qb1**<br>
48 **d6 Qe1+**<br>
49 **Kb3 Qb1+**<br>
50 **Ka4 Qb6**<br>
51 **Qc5 Qa6+**<br>
52 **Kb4 Qb7+**<br>
53 **Kc4 Qa6+**<br>
54 **Kd5 Qa8+**<br>
55 **Qc6 Qa2+**<br>
56 **Kc5 Qa5+**<br>
57 **Qb5 Qa7+**<br>
58 **Qb6 Qa3+**<br>
59 **Kc6 Qa4+**<br>
60 **Kc7 Qc4+**<br>
61 **Qc5 1-0**<br>

Finally convinced that White will be promoting his d-pawn, Black capitulates and respectfully resigns his hopeless position.

<br>

