---
title: 'Game 7: Goldstein''s Revenge'
author: "Jay Shapiro"
output: 
  html_document:
    theme: cerulean
  pdf_document: default
---

<br>

1 **e4 c5**<br>
2 **Nf3 d6**<br>
3 **d4 cxd4**<br>
4 **Nxd4 Nf6**<br>
5 **Nc3 a6**<br>
6 **Be3 e5**<br>
7 **Nb3 Be6**<br>
8 **f3 h5**<br>
9 **Qd2 Nbd7**<br>
10 **0-0-0 b5**<br>
11 **Kb1 Be7**<br>
12 **h3 h4**<br>
13 **Rg1 ...**<br>


```{r echo = FALSE, warning = FALSE, message = FALSE, include = FALSE}
library(rchess)
library(tidyverse)
library(ggplot2)
library(plotly)

game <- Chess$new()

fen <- "r2qk2r/3nbpp1/p2pbn2/1p2p3/4P2p/1NN1BP1P/PPPQ2P1/1K1R1BR1 b kq - 1 13"

game$load(fen)
```
```{r echo = FALSE, warning = FALSE, message = FALSE}

plot(game)

```

<br>

Deviating from all previous games at this juncture; here most peple play f4, which is superior because it contests Black's strong center, threatens a possible f5 thrust, and doesn't spend a turn on moving the rook when the g-file might not be where it needs to be deployed. 

<br>

13 **... Nb6**<br>
14 **g4 d5?**<br>

Black pulls the trigger too early, blowing his chance.