---
title: "Game 4: Shapiro vs Leopold Le Ruyet"
author: "Jay Shapiro"
date: "4/xx/2020"
output: 
  html_document:
    theme: cerulean
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>

You know the drill by now.

<br>

1 **e4 e5**<br>
2 **Nf3 Nc6**<br>
3 **Bb5 a6**<br>
4 **Ba4 Nf6**<br>
5 **0-0 Be7**<br>
6 **Re1 b5**<br>
7 **Bb3 0-0**<br>
8 **c3 d5**<br>

```{r echo = FALSE, warning = FALSE, message = FALSE, include = FALSE}
library(rchess)
library(tidyverse)
library(ggplot2)
library(plotly)

game <- Chess$new()

fen <- "r1bq1rk1/2p1bppp/p1n2n2/1p1pp3/4P3/1BP2N2/PP1P1PPP/RNBQR1K1 w - - 0 9"

game$load(fen)
```
```{r echo = FALSE, warning = FALSE, message = FALSE}

plot(game)

```

<br>

The move 8...d5 signals the Marshall Attack, an enterprising gambit in the Ruy Lopez where Black forfeits a central pawn for a promising kingside attack. The mainline theory runs very deep in the Marshall, with the lichess.org database having a whopping 934 games played from the position after White's sixteenth move!

<br>

9 **exd5 Nxd5**<br>
10 **Nxe5 Nxe5**<br>
11 **Rxe5 c6**<br>
12 **d4 Bd6**<br>
13 **Re1 Qh4**<br>
14 **g3 Qh3**<br>
15 **Be3 Bg4**<br>
16 **Qd3 f5**<br>

```{r echo = FALSE, warning = FALSE, message = FALSE, include = FALSE}
fen <- "r4rk1/6pp/p1pb4/1p1n1p2/3P2b1/1BPQB1Pq/PP3P1P/RN2R1K1 w - - 0 17"

game$load(fen)

```

```{r echo = FALSE, warning = FALSE, message = FALSE}

plot(game)

```

<br>

Finally deviating from the main line, which is 16...Rae8, and choosing a move which is far less popular but still theory

<br>

17 **f4 Rae8**<br>
18 **Nd2 Kh8**<b>
19 **Bxd5 cxd5**<br>
20 **a4 bxa4**<br>
21 **Qxa6 Rxe3!**<br>

This move surprised me; I was expecting the more natural Re6, but both moves are playable and have been played before.

<br>

22 **Rxe3 Bxf4**<br>
23 **Qd3?! ...**<br>

Finally deviating from previously played games (23 Nf1) and striking out into uncharted territory. Although Qd3 is objectively worse than Nf1, I had overweening ambitions about my prospects this game, and I was willing to accept a worse position that allowed me to still push for a win.

<br>

23 **... Bxe3+**<br>
24 **Qxe3 f4**<br>
25 **Qe7 Kg8**<br>
26 **Rf1 f3**<br>
27 **Rf2 h5?**<br>

Black makes the mistake of boxing his own queen into a claustrophobia-inducing prison on h3. Black's best option would have been to extricate his piece with Qh5 and give up on the pipe dream of mating on g2.

<br>

28 **Qe5 h4**<br>
29 **gxh4? ...**<br>

The simple 29 Qxd5+ Kh8 30 Qe5 was superior for White.

<br>

29 **... Rf6?**<br>

Again, the simpler move 29...Qxh4 would have been stronger. Black makes matters needlessly complicated.

<br>

30 **Qg3 ...**<br>

Still ignoring the pawn on d5 in favor of the king safety, and also trying to bait Black into playing that move that he does. I thought it was very likely he would automatically play it and overlook the refutation.

<br>

30 **... Rg6??**<br>
31 **Nxf3 Bxf3??**<br>

I was expecting him to play 31...Bf5